### 2010-06-26 doc/thesis/thesis.tex
got q-q plots in

\documentclass{report}
\usepackage[left=3cm,top=2cm,right=3cm]{geometry}

\usepackage{times}
\usepackage[siunitx]{circuitikz}
%\usetikzlibrary{trees}
%\usetikzlibrary{shapes,snakes}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algorithmic}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{cancel}
\usepackage[colorlinks=true]{hyperref}

\doublespacing

\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\qed}{\\$\Box$}
\newcommand{\qle}{\stackrel{?}{\le}}
\newcommand{\qeq}{\stackrel{?}{=}}
\newcommand{\closure}{\overline}
\newcommand{\intersect}{\cap}
\newcommand{\union}{\cup}
\newcommand{\nullset}{\emptyset}
\newcommand{\minus}{\ \backslash\ }

\begin{comment}

:Author: Micah Chambers
\end{comment}

\begin{document}
\title{BOLD Parameter Estimation Using Particle Filters}
\author{Micah Chambers\\Virginia Polytechnic Institute and State University}
\date{}
\maketitle

\tableofcontents

\include{intro}
\include{background}
\include{particlefilters}
\include{methods}
\include{results}
\include{conclusion}

\bibliographystyle{apalike}
%\bibliographystyle{abbrvnat}
%\bibliographystyle{abbrv}
\bibliography{library}

\end{document}

%trash
%stream of new developments in modeling the BOLD signal, nothing has have been
%a plethora of new studies pushing advancements 
%the choice of analysis tools is still relatively limited. Every widely
%available analysis tool is based on parametric modeling, which requires
%prior knowledge of the noise distribution. Indeed, the linear methods
%usually applied are known not to be robust to non-gaussian, non-white
%noise, which is why so many pre-processing steps are necessary.
%These limitations cast a long shadow over any experiment that claims to be
%able to reject the null hypothesis, no matter the P-value. While obviously
%there is a correlation between the BOLD signal and actual activity, its
%possible the current parametric tests are over-estimating the actual amount
%of activity. Its also possible that the current methods underestimate
%actual activity. In this paper we propose a robust model-based approach to 
%activation detection that makes no assumptions about the underlying noise
%model.