### 2010-07-24 doc/thesis/background.tex
Cleaned up particle filter and background.

\chapter{Methods of Analyzing FMRI}
\label{sec:Prior Works}
Currently, FMRI is used to determine the location of responses
to stimuli. The method of finding activation is discussed in 
\autoref{sec:Statistical Parametric Mapping}. The goal of this 
work is to move away from the question "Is this region active"
and instead ask "What does the activation look like". Answering
this question necessitates more complex models and require
more computation. Already there have been several other
attempts to model the BOLD response; these works will be discussed
in this chapter.

\section{Statistical Parametric Mapping}
\label{sec:Statistical Parametric Mapping}
Although not strictly the same as parameter calculation from 
FMRI, activation detection is similar and worth discussing. Estimation of 
parameters may be considered generalization of the idea 
of activation detection.
Given the popularity of Statistical Parametric Mapping (SPM) 
it is important to draw a distinction between it and the methods proposed
in this work.

\subsection{Classical Activation Detection}
\label{sec:Square Wave}
The most basic method of analyzing FMRI data is by standard T-test
between resting state and active state samples. Simply put, the
mean is calculated separately for non-stimulus and stimulus time intervals.
A classic t-test may then be applied, giving the probability that the
distributions actually have the same mean. Because of the correlated
noise present in FMRI (\autoref{sec:Introduction Noise}), it is necessary
to apply some sort of high-pass filter to the data. Without applying
such a filter, P values must be set extraordinarily high to prevent
false positives \cite{Smith2007}. If there truly is signal
due to stimuli, the distributions will not actually be independent
Gaussians because activation does not fit a square wave 
(\autoref{sec:BOLD Physiology}). For this reason other methods
are often more used, as discussed in 
\autoref{sec:Current Techniques General Linear Model}.

\subsection{Random Field Theory}
\label{sec:RFT}
SPM methods make significant use of T-Tests across large regions;
however, such T-tests work slightly differently than a single
individual test. A t-test with a p-value of $.05$ over 
10,000 voxels, will on average generate 500 false
positives. This is called the multiple comparison problem. Traditional
Bonferroni Correction
deals with this by requiring each test to pass with P
value of $\frac{.05}{10000}$. The probability of 
a single false positive would then $.05$. Unfortunately this leads to unrealistically
low p-values; so low that it would be impossible for any biological system to satisfy. To
compensate, a Gaussian kernel is applied to smooth the image. This has
the benefit of reducing the noise variance and 
decreasing the effective number of independent measurements. Because
the number of independent measurements is smaller, Bonferroni correction
can theoretically be applied with a lower scaling factor than
the original voxel count \cite{Worsley2004}. As side effect of this,
a single voxel activation is virtually impossible to detect.

\subsection{General Linear Model}
\label{sec:Current Techniques General Linear Model}
The most common FMRI analysis technique is SPM, though there
are more advanced versions than the simple square wave
method discussed in \autoref{sec:Square Wave}. Hierarchal
Models are one important improvement that allows researchers
to combine data across multiple runs, patients and stimuli
(see \cite{Hofmann1997} for more on Hierarchical Modeling). 
Hierarchical Models
concatenate all the data into a single dataset, then perform
a linear fit between a design matrix and the data. The design matrix
encapsulates all known experimental factors such as stimuli,
young/old, etc.  The general linear model is defined as:

\begin{equation}
Y(t) = X(t)\beta + \epsilon(t)
\end{equation}

where $Y(t)$ is the smoothed or de-trended time course of measurements,
$X(t)$ is the design matrix, $\beta$ is a column vector of weights,
and $\epsilon$ is the error. Thus for every time, the measurement is
assumed to be a weighted sum of the columns of $X$ plus some error. The calculation
of $\beta$ is then performed using a maximum likelihood or gradient descent search 
to minimize the error.

\begin{figure}
\centering
\includegraphics[scale=.7]{images/HRF}
\caption{Canonical Hemodynamic Response Function}
\label{fig:HRF}
\end{figure}

As mentioned previously, a square wave stimulus 
does not result in a square wave in the activation of brain regions. 
The BOLD signal is in fact a smoothed version of the 
stimuli. As such, when fitting an FMRI time course to the input,
the input ($X$'s columns) is usually smoothed to reduce bias error. 
The best method, that maintains a linear fit, is convolving the input 
with a Hemodynamic 
Response Function (HRF). The \emph{Hemodynamic Response Function}
mimics the basic shape of BOLD activation, including a delay
due to rise time and fall time. The fitting 
process is then a least squares fit over the space of the vector $\beta$. 
Therefore $Y(t)$ is estimated as a linear combination 
of the columns of $X$.

Smoothing the input with a single HRF poses certain problems.
It is well known that different Hemodynamic Response Functions are necessary 
for different regions of the brain. The Canonical HRF that is most used, 
has been optimized for the visual cortex. As \autoref{sec:Post Stimulus Undershoot}
discussed, there are certainly variations in the shape of the
BOLD response, over both brain regions and patients. Handwerker et. al. (\cite{Handwerker2004})
discuessed the implications of choosing an incorrect HRF, the most important
of which is a definite increase in false negatives. While an atlas
of Hemodynamic Response Functions for each region could definitely 
mitigate this risk, it does not account for variation between patients.
Thus, the inability to fit parameters other than scale hinders 
analysis of activation.

\subsection{Hierarchical Linear Models}
As mentioned previously and discussed extensively
in \cite{Friston2002} and \cite{Hofmann1997}, hierarchical models may be 
applied to account for systematic differences between subjects. For instance, if the
study mixes young and old, incorporating
that into the model is wise, regardless of whether it is the goal of the test. 
The reason to do this is to account for additional variance that may not
be otherwise explainable.  The Hierarchical form used by
\cite{Friston2002} is shown in \autoref{eq:Hierarchical}.

\begin{eqnarray}
\label{eq:Hierarchical}
Y(t) = X_1(t)\theta_1 + \epsilon_1(t)            \nonumber \\
\theta_1(t) = X_2(t)\theta_2 + \epsilon_2(t)     \nonumber \\
...                                              \nonumber \\
\theta_{n-1}(t) = X_n(t)\theta_n + \epsilon_n(t) 
\end{eqnarray}

The Empirical Bayes algorithm is used in both in both \cite{Friston2002} 
and \cite{Hofmann1997}. Note that in Empirical Bayes point estimators 
are used for each $\theta$, rather than full distributions.

\subsection{Discussion}
\label{sec:BackgroundConclusion}
In all, the GLM is useful for determining linear 
dependence of a set of regressors on the output. Unfortunately, as discussed in
\autoref{sec:BOLD Analysis} there are significant nonlinearities 
that almost certainly cause false negatives in the Statistical Parametric
Maps. Unfortunately nonlinear analyses have only recently become feasible,
so the scale of the problem is still unknown. The problem is 
highlighted by the common scenario where no significant
activation can be found in a single run \cite{Riera2004}
\cite{Johnston2008}.

The static nature of the  linear model also limits its potential use. 
Besides not allowing HRF differences between patients, there is no
reasonable way to incorporate other forms of physiological
data. Combined FMRI CBF or CBV imaging methods are improving,
as seen in \cite{Chen2009}. These techniques could shed light on
neural activation by providing extra measurements, yet a 
physiologically reasonable model is necessary to incorporate this extra data.
Activation detection methods also don't have the ability 
to identify pathologies based on state variables or parameters. For
example, decreased compliance of
blood vessels could indicate, or even cause, a neurological condition that 
is not easily seen in other imaging modalities.

\section{Approaches to the Balloon Model}
Unlike Statistical Parametric Mapping, the techniques described in this
section are all attempts to learn some version of the BOLD model. 
Although \cite{Buxton1998} and \cite{Friston2000}
both proposed physiologically reasonable values for the model parameters, 
\cite{Friston2002b} was the first paper to calculate the parameters based 
on actual FMRI data. However, in that case, the voxels were chosen from
regions that were detected as active by the GLM. It is therefore possible
that the parameters are biased toward parameters that fit the linear model.

\subsection{Linear Approximation}
In \cite{Friston2002b}, a novel combination of linear and nonlinear modeling
was used to generate parameter estimates. Because there is no closed 
form solution to the ballon model, it impossible to 
calculate the Jacobian matrix $\frac{\partial Y}{\theta}$. Thus 
\cite{Friston2002b} approximates the partial using a Volterra Kernel. 
At each step of the Expectation-
Maximization (EM) algorithm, the differential equation is integrated, calculating
the residuals. Then, for each $\theta_i$ surrounding the current estimate of 
$\theta$, a Volterra-Kernel expansion of the output $y$ is generated. Generation
of the Volterra Kernel is quick, and, since there
is an analytical solution calculating partial derivatives is easy. 
 The full E-M algorithm
for estimating the states is notation-heavy and can be found in \cite{Friston2002b}.

There are a few caveats with this method of optimization. First, the 
partials are based on an approximate values (using Volterra-Kernels) of $y$.
Importantly, the Volterra-Expansion of $y$ is not able to model interactions
between state variables (\cite{Friston2002b}); which certainly increases error.
Additionally, all the tests performed
in \cite{Friston2002b} were on regions found to be active by the General
Linear Model. For this reason, the reliability of the approximation is unknown
for regions that are active but
sufficiently nonlinear to avoid detection by conventional tests.

\subsection{Nonlinear Least Squares}
\label{sec:Nonlinear Least Squares}
Although there are certainly benefits to using a derived model, rather
than a purely empirical model, there are serious implications. The
first problem is that all the powerful classical techniques of gradient
descent are off limits; since the model is a true nonlinear dynamical
system with no closed form solution (although \autoref{sec:Linear Approximation}
circumvented this by calculating a Volterra Series approximation). Without
a Jacobian for residuals, the Gauss-Newton method
is impossible. Additionally, a gradient descent is slow
without the ability to form partials of the output with respect
to all the parameters.

Still, there are other heuristic algorithms
that could illuminate the BOLD response. 
Simulated Annealing (SA) is a common method of optimizing high dimensional
regression problems. First the program selects a random starting point, then
at each iteration it selects a random point in the neighborhood. If the 
that new point is 
below some energy constraint (energy is a function of the residual), 
called the temperature, the algorithm moves
to that point and continues with the next iteration. The temperature
is slowly lowered until no nearby points below the temperature can
be found. There are
variations of this, for instance it is common to require every movement
to be in the downward direction (in terms of energy). Like most nonlinear
optimization problems, there is no guarantee of an optimal solution,
although the longer the algorithm is allowed to run, the better the solution.
Since every step requires a re-integration of the  
balloon model, it can be extremely time consuming, which is why I did 
not use it here.

\begin{algorithm}
\caption{Simulated Annealing Algorithm}
\label{alg:Simulated Annealing}
\begin{algorithmic}
\STATE Initialize $\Theta$, or if there exists a decent estimate start there
\STATE Initialize temperature, T to value above initial energy
\WHILE{$E(\Theta) < T$}
    \REPEAT
        \STATE Pick $\theta$ near $\Theta$
        \STATE Calculate energy, $E$, of $\theta$
    \UNTIL{$E > T$}
    \STATE Move to new estimate: set $\Theta = \theta$
\ENDWHILE
\end{algorithmic}
\end{algorithm}

%[simulated annealing image?] todo

Genetic algorithms (GA) are similar to Simulated Annealing, in
that they randomly move to better solutions based on a cost function.
However; in genetic algorithms a single point estimate isn't used. Instead
a population of estimates is generated, each with distinct parameters,
and then each set of parameters is rated with a fitness function. Parameter
sets that are good get a higher weight. New parameter sets are generated by 
randomly combining pieces of the old parameter sets. The pieces are 
chosen at a rate proportional to the fitness of the donor; thus good
parameter sets tend to pass on their properties. In addition to this,
random mutations are introduced that come from no existing parent. 
The fitness function is then used to weight the new generation, and
the entire process starts over. The stop condition for a genetic algorithm
is typically based on some threshold for fitness or a maximum number 
of generations. As with simulated annealing (or any generic non-linear optimization),
there is no guarantee that a global minimum has been reached.
%[genetic algorithm picture] todo

\begin{algorithm}
\caption{Genetic Algorithm}
\label{alg:Genetic Algorithm}
\begin{algorithmic}
\STATE Initialize $N$ estimates, $E = \{\Theta_0, \Theta_1, ... \Theta_N\}$
\FOR{G generations}
    \STATE Calculate fitness for each $\Theta$, Ex. for residual $R$, $1/R$ or, $e^{-R}$
    \FOR{$i$ in $N$}
        \STATE Randomly select two parents (with higher probability for more fit $\Theta$'s)
        \STATE Randomly merge parts of the two parents to form a new $\Theta_i$
        \STATE With low probability introduce random mutations to parameters in $\Theta_i$
    \ENDFOR
\ENDFOR
\end{algorithmic}
\end{algorithm}

Although both these methods can be highly effective, they have the downside of
requiring high computation time. In the case of the BOLD model,
each time the energy or fitness needs to be calculated, a large number of cycles
must be spent re-simulating the BOLD model for the set of parameters. As I'll
discuss in \autoref{sec:Particle Filter}, the Particle Filter method is able
to circumvent this re-calculation to some degree.

\subsection{Unscented Kalman Filter}
\label{sec:Unscented Kalman Filter}
The Unscented Kalman Filter (UKF) is a powerful Gaussian/Bayes filter that attempts
to model the posterior distribution of dynamical systems as a multivariate
Gaussian. The Unscented Kalman Filter (UKF) generalizes the Extended Kalman
Filter by allowing the state update to be a function, $g$,

\begin{eqnarray}
X(t) &=& g(u(t), X(t-1))\\
Y(t) &=& h(X(t)
\end{eqnarray}

In order to estimate the posterior at $t$, a deterministic set of sigma points 
(often 2 per dimension, plus 1 at the mode of the multivariate distribution)
weighted according to a Gaussian estimate of $X(t-1)$ are passed through
the update equation. This set of points are then used to estimate the 
mean and covariance of $X(t)$. The benefit of this, is that it requires
no Jacobian and only a few extra calculations to get a decent estimate of
a posterior Gaussian. Hu et. al. used the UKF to  perform a similar type of analysis to
the one performed in this work (\cite{Hu2009}).

\begin{figure}
\includegraphics[width=16cm]{images/kalman}
%\includegraphics[trim=6cm .75cm 6cm .75cm,width=16cm]{images/gauss_step_point1sec_3sigma.pdf}
\caption{Example updates of a distribution using Kalman Filter, \cite{Thrun2005}}
\label{fig:EKFWorking}
\end{figure}

The difficulty
of using a Kalman Filter, however, is that it assumes a multivariate 
Gaussian for the state variables, $X(t-1)$. The more nonlinear the system
gets the more likely that the Gaussian will be insufficient to describe
the distribution, $X(t)$. When this occurs, every step from  $X(t+1)$ 
to $X(t)$ will introduce additional
error in the posterior distribution. Furthermore, it is not really known what 
sort of underlying distributions may exist in such a mixed biological,
mechanical, chemical system such as the brain. Assuming the parameters
all to be Gaussian could be a gross error. On the other hand, for
small variances and short time steps the Gaussian distribution is a good 
fit, and so in some cases the Unscented Kalman Filter could work quite
well. The assumption of Gaussianity is what allows
the UKF to estimate the posterior using only the first and second moments;
however, if this assumption is violated significant error will result.

\begin{table}[t]
\centering
\begin{tabular}{|c || c |}
\hline 
Parameter & Run 1 \\
\hline
$\tau_0$ & .98  \\
$\alpha$ & .33 \\
$E_0$ & .34  \\
$V_0$ & .03  \\
$\tau_s$ & 1.54  \\
$\tau_f$ & 2.46  \\
$\epsilon$ & .54  \\
$V_t$ & N(1, .09)  \\
$Q_t$ & N(1, .09)  \\
$S_t$ & N(1, .09) \\
$F_t$ & N(1, .09) \\
\hline
\end{tabular}
\caption{Parameters used to test Gaussianity of variables after being transitioned through
the BOLD model}
\label{tab:steptable} 
\end{table}

To determine the amount of error incurred in a Gaussian estimate during
a typical sample period, I assigned the states of BOLD equations according
to a four dimensional Gaussian. I then propagated the states through 
two seconds of simulation (a typical TR in FMRI) and plotted the resulting
marginal distributions against a Gaussian distribution. This also demonstrates
the degree of nonlinearity present in the system. The parameters used are 
shown in \autoref{tab:steptable}

In order to drive the system without input I intentionally set $s_t$ a non-steady
state, but physiologically plausible, value. The value of $u$ is left at zero 
the entire time, so the 
system would decay naturally (see \autoref{sec:BOLD Physiology})
\autoref{fig:transp1s} shows the results when the system after the
system ran for 100 milliseconds. 
The Q-Q plots fit will with a Gaussian, demonstrating that at this short
time interval nonlinearities have not yet begun to effect the distribution.
However, \autoref{fig:trans1s} shows the result after 1 second, which is faster
than most FMRI scanners are capable of sampling at. At that range the tails 
of the distributions for $v$ and $q$ started to deviate from the
Gaussian distribution. As a result the uncertainty in $y$ deviated from
the Gaussian distribution as well. This is important, because although 
approximating the distribution with a Gaussian based on the first two
moments will worked in the short run, within a period less than typical
measurement rates the distribution deviated from a Gaussian substantially.

Thus, even without introducing variation in the model parameters,
a distinct nonlinearity and non-Gaussianity was present. 
More advanced tests where parameters (especially $\alpha$) are varied
would certainly introduce more error into the Gaussian estimate. 
For this reason, estimating the posterior distribution using only 
the first two moments, which is the basis for the UKF, is not wise.

\begin{figure}
\centering
\includegraphics[trim=6cm .75cm 6cm .75cm,width=16cm]{images/gauss_step_point1sec_3sigma.pdf}
\caption{Distributions of state variables after simulating for .1s}
\label{fig:transp1s}
\end{figure}

\begin{figure}
\includegraphics[trim=6cm .75cm 6cm .75cm,width=16cm]{images/gauss_step_1sec_3sigma.pdf}
\caption{Distributions of state variables after simulating for 1s}
\label{fig:trans1s}
\end{figure}

\subsection{Hybrid Methods}
In \cite{Riera2003}, a maximum
likelihood method for innovation processes was used, as described by
\cite{Ozaki1994}. Ozaki et. al. used a similar construction to a 
Kalman filter to filter the BOLD signal. 
The method used in \cite{Riera2003} performs maximum likelihood on
the innovations rather than the absolute signal levels. 
By using a Local Linearization Filter,
innovation noise such as noise in $\dot{f}$ is turned into simple 
Gaussian noise. This approach is useful when the DC portion of the signal is
of no use; however, it depends greatly on the type of signal. I found that
this method was not very effective for the BOLD signal, when I used it 
with the particle filter approach discussed in the next chapter.

In \cite{Johnston2008}, a hybrid particle filter/gradient
descent algorithm was used to simultaneously derive the static and dynamic 
parameters, (classically known as parameters and state variables, respectively).
A particle filter is used to calculate the state variables at each
time; then the estimated distribution of the particles is used to find
the most likely set of parameters that would give that distribution of state variables.
This process is repeated until the parameters converge. Interestingly
\cite{Johnston2008} comes to a very different set of parameter estimates as compared
to the original \cite{Friston2000} estimates (\autoref{tab:Params}). In fact the results 
are significantly different from virtually every other study. The most obvious discrepancy
is the larger time constants, $\tau_f$, $\tau_s$ and $\tau_0$. 
While of course this could be poor convergence of the algorithm, there is another other possibility.
Unlike all the other methods mentioned, excepting the methods in \autoref{sec:Nonlinear Least Squares},
the algorithm described in \cite{Johnston2008} does not depend on prior distributions.
It is possible then that the bias toward the prior in other methods skewed their results. 
While \cite{Johnston2008} is certainly in the minority; further exhaustive studies 
of the parameters, using unbiased techniques may be called for. A further 
comparison between the distributions found in \cite{Johnston2008} and 
\cite{Friston2000} will be discussed in \autoref{sec:PriorDist}, including
graphs: \autoref{fig:MeanResponseF}, \autoref{fig:MeanResponseJ}.

\section{Conclusion}
Currently there is no ideal solution to solving this system of nonlinear
equations. Exhaustive search type methods such as those employed by
\cite{Johnston2008} and \cite{Vakorin2007} have long
run times even for a single voxel. While Volterra models are an interesting
solution,there have not yet been exhaustive tests to determine whether
such approximations work well throughout state space. The most promising
method of those reviewed here is the Kalman filter based method. It is
able to maintain a fast runtime while still approaching the solution.
While the reliance on a Gaussian estimate to the true posterior distribution could
cause problems, some modifications could make it powerful. 
The particle filter method proposed in the next section bears a strong 
resemblance to the unscented Kalman filter; albeit with more points estimating
the posterior.