### 2010-08-15 doc/thesis/thesis.tex

% PROJECT: <ETD> Electronic Thesis and Dissertation Initiative
%   TITLE: LaTeX report template for ETDs in LaTeX
%  AUTHOR: Neill Kipp, nkipp@vt.edu
%     URL: http://etd.vt.edu/latex/
% SAVE AS: etd.tex
% REVISED: September 6, 1997
% 
%\documentclass[12pt,dvips]{report}
\documentclass[12pt]{report}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.5in}
\setlength{\evensidemargin}{0in}
\setlength{\oddsidemargin}{0in}
\setlength{\topmargin}{0in}

\setlength{\parindent}{0pt}
\setlength{\parskip}{0.1in}

%\usepackage[left=3cm,top=2cm,right=3cm]{geometry}

 \usepackage[hang,small,bf]{caption}

\usepackage{colortbl}
\usepackage{times}
\usepackage[siunitx]{circuitikz}
%\usetikzlibrary{trees}
%\usetikzlibrary{shapes,snakes}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{cancel}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue]{hyperref}
\def\chapterautorefname{Chapter}
\def\sectionautorefname{Section}
\def\subsectionautorefname{Section}

\usepackage{algorithm}
\numberwithin{algorithm}{chapter}
\usepackage{algorithmic}

%\doublespacing

\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\qed}{\\$\Box$}
\newcommand{\qle}{\stackrel{?}{\le}}
\newcommand{\qeq}{\stackrel{?}{=}}
\newcommand{\closure}{\overline}
\newcommand{\intersect}{\cap}
\newcommand{\union}{\cup}
\newcommand{\nullset}{\emptyset}
\newcommand{\minus}{\ \backslash\ }

\begin{comment}

:Author: Micah Chambers
\end{comment}

\begin{document}
\thispagestyle{empty}
\pagenumbering{roman}
\begin{center}

% TITLE
{\Large 
Full Brain Blood-Oxygen-Level-Dependent Signal Parameter Estimation Using Particle Filters
}

\vfill

Micah C. Chambers

\vfill

Thesis submitted to the Faculty of the \\
Virginia Polytechnic Institute and State University \\
in partial fulfillment of the requirements for the degree of

\vfill

Master of Science \\
in \\
Electrical Engineering

\vfill

Chris L. Wyatt, Chair \\
William T. Baumann\\
Aloysius. A. Beex \\
Daniel J. Stilwell

\vfill

August 30, 2010\\
Blacksburg, Virginia

\vfill

Keywords: BOLD Response, FMRI, Nonlinear Systems, Particle Filter, Bayesian Statistics, System Identification
\\
Copyright 2010, Micah C. Chambers

\end{center}

\pagebreak

\thispagestyle{empty}
\begin{center}

{\large Full Brain Blood-Oxygen-Level-Dependent Signal Parameter Estimation Using Particle Filters}

\vfill

Micah C. Chambers

\vfill

(ABSTRACT)

\vfill

\end{center}

Traditional methods of analyzing FMRI images use a linear combination of
just a few static regressors. This works demonstrates an alternative
approach using a physiologically inspired nonlinear model. By using a 
particle filter to optimize the model parameters, the computation time
is kept below a minute per voxel without requiring a linearization 
of the noise in the state
variables. The activation results show regions similar those found in 
SPM; however, there are some notable regions not detected by 
SPM. Though the parameters selected by the particle filter based approach
are more than sufficient to predict the BOLD response,
more model constraints are needed to uniquely identify a single set
of parameters. This ill-posed nature explains the large discrepancies
found in other research that attempted to characterize the model parameters.
For this reason the final distribution of parameters is more medically relevant
than a single estimate. Because the output of the particle filter is 
a full posterior probability, the reliance on the mean to estimate 
parameters is unnecessary. This work presents
not just a viable alternative to the traditional method of detecting
activation, but an extensible technique of estimating the joint Probability
Distribution Function of the BOLD parameters.

\vfill

% GRANT INFORMATION

%That this work received support from the Southeastern Universities
%Research Association (SURA) ``Monticello Library Project'' is purely
%coincidental.

\pagebreak

% Dedication and Acknowledgments are both optional
% \chapter*{Dedication}
\chapter*{Acknowledgments}

This thesis would not have been possible without the guidance, advice and
patience of my advisor, Chris Wyatt. I knew nothing about being a researcher
or about medical imaging when I came to him, and he taught me to read broadly
and to always dig deeper. Thank you Dr. Wyatt.

I would also like to thank William Baumann for his help and inspiration. I
have learned an incredible amount reading papers with you and Dr. Wyatt.

To my parents, for all the psychological (and financial) help along the way,
thank you so much. When things were the darkest you were there for me, I couldn't
ask for better parents. Love you guys.

Finally I would like to thank my friends, for the support, for the prayer,
and, yes, for the occasional stress-relieving game of Halo, SOASE, and every
type of SC.  You guys are the best.

\tableofcontents
\pagebreak

\listoffigures
\pagebreak

\listoftables
\pagebreak

\pagenumbering{arabic}
\pagestyle{myheadings}

\include{intro}
\include{background}
\include{particlefilters}
\include{methods}
\include{results}
\include{resultsB}
\include{discussion}
\include{conclusion}

\bibliographystyle{plain}
%\bibliographystyle{abbrvnat}
%\bibliographystyle{apalike}
%\bibliographystyle{abbrvnat}
%\bibliographystyle{abbrv}
\bibliography{library}

\end{document}

%\chapter{Introduction}
%\markright{Albert J. Kippleby \hfill Chapter 1. Introduction \hfill}
%
%William Shakespeare has profoundly affected the field of literature
%worldwide.  In the United States there was a surge of Shakespearean
%literature starting in the 1960s, with the opening of the Montgomery
%Shakespearean festival and continuing into the present ...

%%%%%%%%%%%%%%%%%
%
% Include an EPS figure with this command:
%   \epsffile{filename.eps}
%

%%%%%%%%%%%%%%%%
%
% Do tables like this:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% PROJECT: <ETD> Electronic Thesis and Dissertation Initiative
%%   TITLE: LaTeX report template for ETDs in LaTeX
%%  AUTHOR: Neill Kipp, nkipp@vt.edu
%%     URL: http://etd.vt.edu/latex/
%% SAVE AS: etd.tex
%% REVISED: September 6, 1997
%% 
%
%% Instructions: Remove the data from this document and replace it with your own,
%% keeping the style and formatting information intact.  More instructions
%% appear on the Web site listed above.
%
%\documentclass[12pt,dvips]{report}
%
%\setlength{\textwidth}{6.5in}
%\setlength{\textheight}{8.5in}
%\setlength{\evensidemargin}{0in}
%\setlength{\oddsidemargin}{0in}
%\setlength{\topmargin}{0in}
%
%\setlength{\parindent}{0pt}
%\setlength{\parskip}{0.1in}
%
%% Uncomment for double-spaced document.
%% \renewcommand{\baselinestretch}{2}
%
%% \usepackage{epsf}
%
%\begin{document}
%
%\thispagestyle{empty}
%\pagenumbering{roman}
%\begin{center}
%
%% TITLE
%{\Large 
%Use of Metaphor in Shakespeare's Plays and its Potential
%Application in Twenty-first Century Literature
%}
%
%\vfill
%
%Albert J. Kippleby
%
%\vfill
%
%Dissertation submitted to the Faculty of the \\
%Virginia Polytechnic Institute and State University \\
%in partial fulfillment of the requirements for the degree of
%
%\vfill
%
%Doctor of Philosophy \\
%in \\
%Literature and Technology
%
%\vfill
%
%Neill A. Kipp, Chair \\
%Emilio J. Arce \\
%Scott A. Guyer \\
%Laura Weiss
%
%\vfill
%
%July 16, 1997 \\
%Blacksburg, Virginia
%
%\vfill
%
%Keywords: Metaphysics, Information Retrieval, Spacecraft
%\\
%Copyright 1997, Albert J. Kippleby
%
%\end{center}
%
%\pagebreak
%
%\thispagestyle{empty}
%\begin{center}
%
%{\large Use of Metaphor in Shekespeare's Plays and its Potential
%Application in Twenty-first Century Literature}
%
%\vfill
%
%Albert J. Kippleby
%
%\vfill
%
%(ABSTRACT)
%
%\vfill
%
%\end{center}
%
%The need for concrete examples increases when technology becomes
%difficult to explain.  In documentation for computer systems
%especially, we see a wide audience of field experts attempting to
%comprehend documentation for computer software and hardware of which
%they should only require a cursory understanding.  Additionally, as
%the pace of the information age quickens we see document authors
%struggle for \textit{examplia-concretes} with wide applicability, and
%consistently rely on excerpts from Shakespearean literature as a
%public-domain source for their various explications.
%
%We predict the twenty-first century will be no different.  Actuarial
%studies show explosion in the information industry such that four out
%of five persons will be \textit{bona fide} electronic document
%authors; many of those will have one or more college degrees.  We
%prove through computer simulation \textsc{Machinum Simitatores} that
%authors of twenty-first century literature will be affected by these
%examples and will include metaphor with Shakespearean source into
%their writing with increasing frequency.
%
%\vfill
%
%% GRANT INFORMATION
%
%That this work received support from the Southeastern Universities
%Research Association (SURA) ``Monticello Library Project'' is purely
%coincidental.
%
%\pagebreak
%
%% Dedication and Acknowledgments are both optional
%% \chapter*{Dedication}
%% \chapter*{Acknowledgments}
%
%\tableofcontents
%\pagebreak
%
%\listoffigures
%\pagebreak
%
%\listoftables
%\pagebreak
%
%\pagenumbering{arabic}
%\pagestyle{myheadings}
%
%\chapter{Introduction}
%\markright{Albert J. Kippleby \hfill Chapter 1. Introduction \hfill}
%
%William Shakespeare has profoundly affected the field of literature
%worldwide.  In the United States there was a surge of Shakespearean
%literature starting in the 1960s, with the opening of the Montgomery
%Shakespearean festival and continuing into the present ...
%
%%%%%%%%%%%%%%%%%%
%%
%% Include an EPS figure with this command:
%%   \epsffile{filename.eps}
%%
%
%%%%%%%%%%%%%%%%%
%%
%% Do tables like this:
%
% \begin{table}
% \caption{The Graduate School wants captions above the tables.}
%\begin{center}
% \begin{tabular}{ccc}
% x & 1 & 2 \\ \hline
% 1 & 1 & 2 \\
% 2 & 2 & 4 \\ \hline
% \end{tabular}
%\end{center}
% \end{table}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%% If you are using BibTeX, uncomment the following:
%% \thebibliography
%%
%% Otherwise, uncomment the following:
%% \chapter*{Bibliography}
%
%% \appendix
%
%% In LaTeX, each appendix is a "chapter"
%% \chapter{Program Source}
%
%
%% Finally, the VITA
%\chapter*{Vita}
%
%Albert was born on a sunny day...
%
%\end{document}